---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# wavdrcast

<!-- badges: start -->
<!-- badges: end -->

This package provides forecasts of time series using a
direct multistep model where regressors can be composed by
wavelet-based signal estimation. A functional available in  **wavdrcast** allows one to
choose a good wavelet specification for sinal extraction
considering  the forecast
criterion.  The scope is not limited to wavelet regressors,
however. In general, any other data set can be used as
an exogenous variable, including AR components of the time series. This can be
interesting when one is comparing the forecast property of
core inflation measure or output gap estimators.

## Installation

You can install  the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("wavdrcast")
devtools::install_github("nelson16silva/wavdrcast", build_vignettes = TRUE)
```

## Example


```{r example}
library(wavdrcast)
```

See **wdrcast**'s vignette:

```{r, eval = FALSE}
vignette("intro", package = "wavdrcast")
browseVignettes("wavdrcast")
```



